---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stoppingrule

<!-- badges: start -->
<!-- badges: end -->

## R package "stoppingrule": Create and describe early stopping rules for clinical studies

stoppingrule is an R package that provides functionality for constructing,
describing, and evaluating stopping rules. Clinical trials often include these
rules in order to ensure the safety of study treatments and feasibility of the
study.

## Installation

The current version of stoppingrule is available in this repository. If you have
the R package "devtools" installed, stoppingrule can be installed directly from
GitHub with:
``` r
require(devtools)
install_github("mjmartens/stoppingrule")
```

## Example

Suppose a clinical trial of 30 patients is being developed. Study investigators
wish to monitor the occurrence of a specific toxicity type, which occurs in 20%
of patients in the target population based on historical data. We want to create
a stopping rule to test whether the rate of a given toxicity exceeds 20% using a
type I error rate of 10%. Moreover, we wish to check the stopping rule
continuously, after each patient completes follow-up for the toxicity endpoint.

The function `calc.rule.bin` creates a 'rule.bin' object which contains a matrix
with the numbers of evaluable patients and their corresponding stopping boundary
values as well as the rule's design parameters:

```{r calc}
require(stoppingrule)
bb_rule = calc.rule.bin(ns=1:30,p0=0.20,alpha=0.10,type="BB",param=c(0.6,2.4))
print(bb_rule)
```

The function call uses the Bayesian beta-binomial model approach proposed by 
Geller et al. 2003 to construct the stopping boundaries and displays the
boundaries for the first few patients. A weakly informative Beta(0.6,2.4) is 
specified for the toxicity probability. The `bb_rule$Rule' element of the output
displays the number evaluable and stopping criteria for all analyses; other
elements contain the design parameters and boundary parameter for the stopping
rule. For this rule, rejection is impossible with 1 or 2 evaluable patients
because the corresponding boundary values exceeds these numbers.

The function `table.rule.bin` can produce a succinct summary of the stopping
rule for the entire cohort from the rule calculated above:

```{r table}
table.rule.bin(bb_rule)
```

We can also obtain a graphical summary of the stopping rule using the `plot`
function:

```{r plot, echo = FALSE}
plot(bb_rule)
```

Lastly, the `OC.rule.bin` function can assess the operating characteristics of
this stopping rule. The rejection probability and expected numbers of events at
the time of stopping are computed at true toxicity rates of p = 20%, 25%, 30%,
35%, and 40% as follows:

```{r OC}
OC.rule.bin(rule=bb_rule,ps=seq(0.2,0.4,0.05))
```
